### 2022-02-08 ms_body/manuscript.Rmd
Make Figure 7

Ecological models are important tools to support and guide the development and implementation of environmental policies and management programs. Species distributions models (SDM's), such as the popular MaxEnt software, are frequently used to guide conservation programmes, predict the potential distribution of invasive species, forecast the impacts of climate change and develop applied ecological research (e.g. biological control). Many applications of these models require that SDM's are transferred in either space and/or time. However, few studies to date have tested the transferability, and thus usefulness, of MaxEnt models using fully independent data. Moreover, numerous authors have raised concerns over how model complexity (controlled primarily by feature class and regularisation multiplier settings) may affect MaxEnt model transferability. In this paper, we evaluated the usefulness of MaxEnt models transferred in space using a native Australian insect, *Dasineura rubiformis* Kolesik (Diptera: Cecidomyiidae) that has been intentionally introduced as a biological control agent of the invasive plant, *Acacia mearnsii* De Wild (Fabaceae) in South Africa. MaxEnt models were developed using native-range records only (Australia) and projected over South Africa to identify the potential climatic suitability for the insect, using a range of model settings configurations. Model transferability was assessed using independent post-release data from South Africa. Our results demonstrated that MaxEnt scores were positively correlated with increased establishment rates for *D. rubiformis* in South Africa. However, we found that this pattern held only when MaxEnt models were specified with user-defined settings configurations. A MaxEnt model calibrated with default settings was overfit to the training data, and performed poorly when transfered in space. Our study demonstrates that MaxEnt models can prove useful when transferred in space, however, user-optimised model settings likely need to be used to control model complexity, and reduce the likelihood of overfitting.

@@

Spatial autocorrelation is an important factor that may affect model outputs. Filtering of species occurrence data may limit the inherent biases in the data and improve model quality [@Veloz2009]. To avoid pseudo-replication, only one occurrence record per 2.5 minute grid cell was used for model calibration. Species occurrence datasets were thinned using the ‘spThin’ package [@Aiello-Lammens2015], and spatial autocorrelation analyses were performed using the ‘ecospat’ package [@DiCola2017].

@@

Our results indicated that MaxEnt SDM's were able to provide robust estimates of climatic suitability for *D. rubiformis* in South Africa. AUC values greater than 0.75 are typically considered good fitting models (having high predictive power) [@Fielding1997]. When MaxEnt models, calibrated using native-range records for *D. rubiformis* from Australia, were projected over South Africa, they had AUC scores of 0.83 - 0.84, indicative of high power to discriminate between sites where *D. rubiformis* had established in South Africa or not. Additionally, we showed that the probability that *D. rubiformis* was recorded at a field site in South Africa increased with higher MaxEnt climatic suitability scores. This result demonstrates the usefulness of the MaxEnt SDM and that, at least for our case-study, SDM's can be transferred into novel geographic regions, and provide valuable information to guide policy making and the development of management programmes. For example, biological control programmes such as the programme in which *D. rubiformis* is used to curtail the spread of the highly invasive weed *A. mearnsii* expend many resources (time and money) to mass-rear insect populations that are then released into the field. The positive correlation between MaxEnt suitability scores and *D. rubiformis* establishment means that researchers can develop and implement mass-rearing and experimental release programmes for the insect that prioritise releasing the insect at sites that are highly climatically suitable, and thus, more likely the insect will establish viable populations. This may assist in improving the level of control over the target weed and increasing the efficiency of resources expended by increasing the chances of establishing viable insect populations.

Our study adds to a growing body of literature indicating that SDM model complexity affects model performance and transferability when being projected into novel geographic regions. While the default MaxEnt model had the highest predictive power in the training area (Australia), this model performed extremely poorly when transferred over South Africa. In contrast, the four MaxEnt models calibrated with tuned settings showed high predictive power when transferred. The poor transferability of the Default model is likely explained by default settings overfitting the model on the training points [@Warren2014]. This finding corroborates other studies showing that high predictive power in the model training area is not a guarantee of transferability [@Warren2014; @Duque-Lazo2016]. We echo that recommendations of @Shcheglovitova2013 and @Muscarella2014 that model tuning should be performed to estimate optimal model complexity when transferring MaxEnt models in space and/or time in order to maximise their usefulness.

@@

Given that the *D. rubiformis* MaxEnt models showed high predictive accuracy when transferred in space over South Africa and the statistically significant positive correlation between MaxEnt scores and *D. rubiformis* establishment, we wish to highlight how these models can be used in an applied setting. To visualise the potential geographic regions where *D. rubiformis* may establish in South Africa, we thresholded the MaxEnt map that optimised AICc where the probability of establishment was > 50% (i.e. MaxEnt score = 0.30), given in Fig. 9. The map demonstrates that the entire south and east coast regions of South Africa are climatically suitable for *D. rubiformis*. Mass-releasing and releasing weed biological control agents is a time-consuming and costly endeavor (Coetzee et al., 2021). As such, our models provide a statistically validated approach to optimising the redistribution of *D. rubiformis* in South Africa by prioritising releases at field sites where MaxEnt scores are > 0.30, and therefore where the probability of establishment was > 50%.

@@

Species distribution models are becoming increasingly popular with researchers, stakeholders, policy makers and land managers alike, particularly for making predictions in novel geographic geographic regions or forecasting species responses to climate change. Transfering SDM's in space and/or time, however, has been met with justifiable resilience and concerns over how well these models perform upon transfer. In this study, we presented a case-study demonstrating that MaxEnt SDM's were able to accurately predict the distribution of an Australian insect when introduced into South Africa. Our study is one of only a few studies to date validating MaxEnt models using spatially independent data. However, the accuracy of the MaxEnt models was dependent on the settings used to calibrate the models, with default MaxEnt programme settings having low predictive power. Our results demonstrate that MaxEnt can be useful in predicting where a species may establish, even when projected into a novel geographic area, but that users need to be aware of how model settings can influence model outputs and usefulness. Demonstrating how SDM's can be useful, while communicating uncertainty in their predictions and outputs, is essential for building confidence in the use of SDM's (Blasi et al. 2020). We showed how validating predictive models can be used to guide the implementation of a biological control programme by optimising the redistribution and release of *D. rubiformis* in geographic areas that are climatically suitable and where the probability of the insect establishing is greater than 50%.

@@

```{r import_fig_1, echo = FALSE}
# knitr::include_graphics(here::here("./ms_body/ms_figs/fig_1.png"), dpi = NA)
```

@@

```{r import_fig_2, echo = FALSE}
# knitr::include_graphics(here::here("./ms_body/ms_figs/fig_2.png"), dpi = NA)
```

@@

```{r import_fig_3, echo = FALSE}
# knitr::include_graphics(here::here("./ms_body/ms_figs/fig_3.png"), dpi = NA)
```

@@

```{r import_fig_4, echo = FALSE}
# knitr::include_graphics(here::here("./ms_body/ms_figs/fig_4.png"), dpi = NA)
```

@@

```{r import_fig_5, echo = FALSE}
# knitr::include_graphics(here::here("./ms_body/ms_figs/fig_5.png"), dpi = NA)
```

@@

```{r import_fig_6, echo = FALSE}
# knitr::include_graphics(here::here("./ms_body/ms_figs/fig_6.png"), dpi = NA)
```

@@

# Add footnote
table1 <- table1 %>%
    footnote(
    alphabet = c("Adjusted to indicate change in Y for each 1% unit increase in covariate"),
    escape = TRUE,
    footnote_as_chunk = FALSE,
    threeparttable = TRUE) 
table1
```

@@

# Create Table S1
tableS1 <- tableS1_data %>%
  knitr::kable(
    format = "latex",
    align = "l",
    booktabs = TRUE,
    longtable = TRUE,
    linesep = "",
    escape = FALSE,
    col.names = c(
      "Model",
      "RM",
      "FC",
      "AUC\\textsubscript{test}",
      "AUC\\textsubscript{diff}",
      "OR\\textsubscript{10}"
    )
  ) %>%
  kableExtra::kable_styling(
    position = "left",
    latex_options = c("striped", "repeat_header")
    #,    stripe_color = "gray!15"
  )

>>>>

Ecological models are important tools to support and guide the development and implementation of environmental policies and management programs. Species distributions models (SDM's), such as the popular MaxEnt software, are frequently used to guide conservation programmes, predict the potential distribution of invasive species, forecast the impacts of climate change and develop applied ecological research (e.g. biological control). Many applications of these models require that SDM's are transferred in either space and/or time. However, few studies to date have tested the transferability, and thus usefulness, of MaxEnt models using fully independent data. Moreover, numerous authors have raised concerns over how model complexity (controlled primarily by feature class and regularisation multiplier settings) may affect MaxEnt model transferability. In this paper, we evaluated the usefulness of MaxEnt models transferred in space using a native Australian insect, *Dasineura rubiformis* Kolesik (Diptera: Cecidomyiidae) that has been intentionally introduced as a biological control agent of the invasive plant, *Acacia mearnsii* De Wild (Fabaceae) in South Africa. MaxEnt models were developed using native-range records only (Australia) and projected over South Africa to identify the potential climatic suitability for the insect, using a range of model settings configurations. Model transferability was assessed using independent post-release data from South Africa. Our results demonstrated that MaxEnt scores were positively correlated with increased establishment rates for *D. rubiformis* in South Africa. MaxEnt model outputs and projections were relatively consistent, irrespective of the settings used. Despite MaxEnt having to extrapolate over much of South Africa, MaxEnt performance was higher than modelling climatic suitability as a function of whether models were in extrapolation or interpolation. Our study demonstrates that MaxEnt models can prove useful when transferred in space, however, users need to be aware of the effect of different model settings configurations and model extrapolation can have on model transferability and usefulness in applied ecological settings.

@@

Spatial autocorrelation is an important factor that may affect model outputs. Filtering of species occurrence data may limit the inherent biases in the data and improve model quality [@Veloz2009]. To avoid pseudo-replication, only one occurrence record per 2.5 minute grid cell was used for model calibration. Species occurrence datasets were thinned using the `spThin` package [@Aiello-Lammens2015], and spatial autocorrelation analyses were performed using the `ecospat` package [@DiCola2017].

@@

Our results indicated that MaxEnt SDM's were able to provide robust estimates of climatic suitability for *D. rubiformis* in South Africa. AUC values greater than 0.75 are typically considered good fitting models (having high predictive power) [@Fielding1997]. When MaxEnt models, calibrated using native-range records for *D. rubiformis* from Australia, were projected over South Africa, they had AUC scores of 0.89 - 0.92, indicative of high power to discriminate between sites where *D. rubiformis* had established in South Africa or not. Additionally, we showed that the probability that *D. rubiformis* was recorded at a field site in South Africa increased with higher MaxEnt climatic suitability scores. This result demonstrates the usefulness of the MaxEnt SDM and that, at least for our case-study, SDM's can be transferred into novel geographic regions, and provide valuable information to guide policy making and the development of management programmes. For example, biological control programmes such as the programme in which *D. rubiformis* is used to curtail the spread of the highly invasive weed *A. mearnsii* expend many resources (time and money) to mass-rear insect populations that are then released into the field. The positive correlation between MaxEnt suitability scores and *D. rubiformis* establishment means that researchers can develop and implement mass-rearing and experimental release programmes for the insect that prioritise releasing the insect at sites that are highly climatically suitable, and thus, more likely the insect will establish viable populations. This may assist in improving the level of control over the target weed and increasing the efficiency of resources expended by increasing the chances of establishing viable insect populations.

Our study adds to a growing body of literature indicating that SDM model complexity affects model performance and transferability when being projected into novel geographic regions. While the default MaxEnt model had the high predictive power in the training area (Australia), this model was somewhat overfit to the training data [@Warren2014]. This finding corroborates other studies showing that high predictive power in the model training area is not a guarantee of transferability [@Warren2014; @Duque-Lazo2016]. We echo that recommendations of @Shcheglovitova2013 and @Muscarella2014 that model tuning should be performed to estimate optimal model complexity when transferring MaxEnt models in space and/or time in order to maximise their usefulness.

@@

Given that the *D. rubiformis* MaxEnt models showed high predictive accuracy when transferred in space over South Africa and the statistically significant positive correlation between MaxEnt scores and *D. rubiformis* establishment, we wish to highlight how these models can be used in an applied setting. To visualise the potential geographic regions where *D. rubiformis* may establish in South Africa, we thresholded the MaxEnt map that optimised AICc where the probability of establishment was > 75% (i.e. MaxEnt score = 0.30), given in Fig. 7. The map demonstrates that the entire south and east coast regions of South Africa are climatically suitable for *D. rubiformis*. Mass-releasing and releasing weed biological control agents is a time-consuming and costly endeavor (Coetzee et al., 2021). As such, our models provide a statistically validated approach to optimising the redistribution of *D. rubiformis* in South Africa by prioritising releases at field sites where MaxEnt scores are > 0.30, and therefore where the probability of establishment was > 75%.

@@

Species distribution models are becoming increasingly popular with researchers, stakeholders, policy makers and land managers alike, particularly for making predictions in novel geographic geographic regions or forecasting species responses to climate change. Transfering SDM's in space and/or time, however, has been met with justifiable resilience and concerns over how well these models perform upon transfer. In this study, we presented a case-study demonstrating that MaxEnt SDM's were able to accurately predict the distribution of an Australian insect when introduced into South Africa. Our study is one of only a few studies to date validating MaxEnt models using spatially independent data. However, the accuracy of the MaxEnt models and their climatic suitability projections were dependent on the settings used to calibrate the models. Our results demonstrate that MaxEnt can be useful in predicting where a species may establish, even when projected into a novel geographic area, but that users need to be aware of how model settings and model extrapolation can influence model outputs and usefulness. Demonstrating how SDM's can be useful, while communicating uncertainty in their predictions and outputs, is essential for building confidence in the use of SDM's (Blasi et al. 2020). We showed how validating predictive models can be used to guide the implementation of a biological control programme by optimising the redistribution and release of *D. rubiformis* in geographic areas that are climatically suitable and where the probability of the insect establishing is greater than 75%.

@@

```{r import_fig_1, echo = FALSE}
knitr::include_graphics(here::here("./ms_body/ms_figs/fig_1_distributions.png"), dpi = NA)
```

@@

```{r import_fig_2, echo = FALSE}
knitr::include_graphics(here::here("./ms_body/ms_figs/fig_2_model_tuning.png"), dpi = NA)
```

@@

```{r import_fig_3, echo = FALSE}
knitr::include_graphics(here::here("./ms_body/ms_figs/fig_3_logits.png"), dpi = NA)
```

@@

```{r import_fig_4, echo = FALSE}
knitr::include_graphics(here::here("./ms_body/ms_figs/fig_4_auc.png"), dpi = NA)
```

@@

```{r import_fig_5, echo = FALSE}
knitr::include_graphics(here::here("./ms_body/ms_figs/fig_5_maxent_maps.png"), dpi = NA)
```

@@

```{r import_fig_6, echo = FALSE}
knitr::include_graphics(here::here("./ms_body/ms_figs/fig_6_mess.png"), dpi = NA)
```

@@

# Add footnote
table1 <- table1 %>%
    footnote(
    alphabet = c("Adjusted to indicate change in Y for each 1% unit increase in climatic similarity"),
    escape = TRUE,
    footnote_as_chunk = FALSE,
    threeparttable = TRUE) 
table1
```

@@

# Create Table S1
tableS1 <- tableS1_data %>%
  knitr::kable(
    format = "latex",
    align = "l",
    booktabs = TRUE,
    longtable = TRUE,
    linesep = "",
    escape = FALSE,
    col.names = c(
      "Model",
      "RM",
      "FC",
      "AUC\\textsubscript{test}",
      "AUC\\textsubscript{diff}",
      "OR\\textsubscript{10}",
      "AIC\\textsubscript{c}"
    )
  ) %>%
  kableExtra::kable_styling(
    position = "left",
    latex_options = c("striped", "repeat_header")
    #,    stripe_color = "gray!15"
  )
